An Analysis of the Use of Average Centipawn Loss to Detect Cheating in Chess

### 2022-12-02 Manuscript/manuscript.tex
fixed title page and abstract page to follow Dr. Yan example
edits to data

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

@@

%% metadata
\title{An Analysis of the Use of Average Centipawn Loss to Detect Cheating in Chess}
\author{James Horbury\\
    University of Connecticut
}
\date{November 14, 2022}

@@

\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}
\label{sec:abs}

@@

Keywords: Centipawn math, Niemann, Carlsen, Chess

@@

@@

The data used was provided by Lichess.com (via kaggle) \citep{zaidqureshi2022}, a free and open-source internet chess server that holds the records of many professional games. It is a comprehensive list of every professional game Hans Niemann has played, both virtually and OTB, from March of 2018 to September of 2022 and is provided in both .csv and .pkl format. Some of the relevant information it contains are what color Niemann and his opponent played and what their respective ELO scores and calculated ACL scores were along with some other unimportant details for this analysis (e.g., ACL score of Niemann's opponent, tournament name and event round, etc.). The main information of concern with regard to the entirety of the data is Niemann's ELO score and calculated ACL score of each game of each particular point in time, as this is what will be used to plot. All other games that included information where Niemann was not playing were dropped from the original dataset provided by Lichess.

>>>>

\documentclass[12pt, letterpaper, titlepage]{article}
\usepackage[utf8]{inputenc}

@@

%% metadata
\title{An Analysis of the Use of Average Centipawn Loss to Detect Cheating in Chess}

\author{James Horbury\\
    University of Connecticut
}
\date{\today}

@@

\begin{abstract}
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}
\label{sec:abs}

@@

\bigskip
\noindent\sc{Kewords}:
Centipawn math; 
Niemann;
Carlsen;
Chess;
\end{abstract}

@@

How was the topic brought about?

@@

The dataset was compiled by Rafael Vicente Leite, an independent chess researcher and data scientis, immediately following the events of the scandal with the intent of conducting his own investigation of the Niemann controversy. The data itself was sourced from ChessBase, a German chess software company that maintains and sells large-scale databases containing the moves of recorded professional chess games. In addition to just holding data from prior games, however, it also provides some engine analysis of these games using Stockfish. Liete converted this raw data, that is to say the archive of games, from their initial portable game notation (PGN) format to data that was more amenable to statistical analysis.

Specifically, the processed dataset is comprised of ranked games (both virtual and OTB) of the following professional chess players: Hans Niemann, Erigaisi Arjun, Gukesh D, Alireza Firouzja, Vincent Keymer, and Igors Rausis.
% I don't know the exact reasoning for why these players were selected aside from Niemann and Carlsen, the two players who sparked the cheating controversy, and Rausis, a known cheater and perhaps "control" of his analysis.

It is a comprehensive list of every professional game they have played from 2018 to 2022 and is provided in both .csv and .pkl format. Some of the relevant information it contains are what color was played, what their respective ELO scores were at the time, and calculated ACL scores were along with some other unimportant details for this analysis (e.g., ACL score of Niemann's opponent, tournament name and event round, etc.). The main information of concern with regard to the entirety of the data is Niemann's ELO score and calculated ACL score of each game of each particular point in time, as this is what will be used to plot. All other games that included information where Niemann was not playing were dropped from the original dataset provided by Lichess.
